Draft Exploration

### 2020-04-01 draft_exploration.Rmd
created hs_unsigned_full dataset of all players who did not sign out of high school along with all subsequent drafts they were in

# joined draft and war data by player mlb ID, then mutated teamname to condense
# the Angel's team names into a single team

master <- draft %>%
  left_join(war_joined, by = c("mlbid" = "mlb_ID")) %>%
  mutate(teamname = as.factor(case_when(teamname %in% c("California Angels",
                                                        "Anahaim Angels",
                                                        "Los Angeles Angels of Anaheim") ~
                                                          "Los Angeles Angels",
                                        TRUE ~ teamname)))

@@

```{r}
hs_unsigned <- master %>%
  arrange(year) %>%
  group_by(mlbid) %>%
  mutate(draft_numb = row_number()) %>%
  filter(first_name == "Brady" & last_name == "Aiken")
```

>>>>

# joined draft and war data by player mlb ID, then mutated teamname to condense
# the Angel's team names into a single team. I then arranged the data by year,
# grouped by player_id, and mutated a column called draft_mumber based on the
# row number (or occurence of that player's id) in the data. I did this because
# I wanted a column that tells for each draft a player is selected in, that it
# is the nth time that player has been selected. For example, if a player turns
# down the draft out of high school, after a year of junior college, and gets
# drafted after his junior year the draft_numb column should say 1,2,3 for each
# respective draft.

master <- draft %>%
  left_join(war_joined, by = c("mlbid" = "mlb_ID")) %>%
  mutate(teamname = as.factor(case_when(teamname %in% c("California Angels",
                                                        "Anahaim Angels",
                                                        "Los Angeles Angels of Anaheim") ~
                                                          "Los Angeles Angels",
                                        TRUE ~ teamname))) %>%
  arrange(year) %>%
  group_by(player_id) %>%
  mutate(draft_numb = row_number()) %>%
  ungroup()

@@

```{r}

# I made a new dataset where I filtered the master set for only unsigned
# highschool picks. I only selected the player_id column. I wanted to do this so
# that I could have a list of IDs and then rejoin these player IDs using a
# left_join to keep only the data from players who did not sign out of high
# school, but all years of their draft data rather than just the draft they
# turned down out of high school. That is what hs_unsigned_full is.

hs_unsigned_list_ids <- master %>%
  filter(source == "H" & signed == "N") %>%
  select(player_id)

hs_unsigned_full <- hs_unsigned_list_ids %>%
  left_join(master, by = "player_id") %>%
  arrange(player_id, desc(year))
```

```{r}
hs_unsigned_full %>%
  filter(first_name == "Grayson" & last_name == "Byrd")
```
